---
title: "Generation audit: session-level retrieval"
subtitle: "Memory-only gold sessions; does not re-estimate distribution effects"
author: "MemMeetDoc"
date: "`r format(Sys.Date(), '%B %d, %Y')`"
format:
  html:
    toc: true
    toc-depth: 3
    number-sections: true
    code-fold: true
    embed-resources: true
    theme: flatly
editor_options:
  chunk_output_type: console
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(
  echo = TRUE, warning = FALSE, message = FALSE,
  fig.width = 8, fig.height = 5, dpi = 120
)
options(width = 110)
```

This check asks whether session-generation attributes made some gold sessions
harder to retrieve. The outcome is a **strict** evidence hit on that session.
Persona and **per-evidence** in-domain status (this item's category vs.\ \(T_p\))
are controls, not a second test of RQ1b. Incomplete sessions (missing key
information in the audit) cannot receive a strict hit; they stay in the primary
fit and are dropped in a sensitivity refit.

# Setup

```{r libraries}
suppressPackageStartupMessages({
  library(lme4)
  library(emmeans)
  library(readr)
  library(dplyr)
  library(tidyr)
  library(ggplot2)
  library(broom.mixed)
  library(car)
  library(knitr)
})

set.seed(42)
emm_options(pbkrtest.limit = 1e5, lmerTest.limit = 1e5)
glmer_ctrl <- glmerControl(optimizer = "bobyqa", optCtrl = list(maxfun = 3e5))

AGENT_LABELS <- c(
  agent_1 = "Separate Store",
  agent_2 = "MemPalace",
  agent_3 = "Corrective",
  agent_4 = "Unified Store"
)
PERSONA_LABELS <- c(
  persona_1 = "Tech",
  persona_2 = "Business-Tech",
  persona_3 = "Sports"
)
theme_thesis <- function(base_size = 13) {
  theme_minimal(base_size = base_size) +
    theme(
      plot.title    = element_blank(),
      plot.subtitle = element_blank(),
      panel.grid.minor = element_blank(),
      panel.grid.major.x = element_blank(),
      legend.position = "bottom",
      strip.text = element_text(face = "bold")
    )
}

relabel <- function(x, map) factor(unname(map[as.character(x)]), levels = unname(map))
```

```{r paths}
find_repo_root <- function() {
  candidates <- c(".", "..", getwd())
  for (cand in candidates) {
    if (file.exists(file.path(cand, "data", "analysis", "model_features_audit.csv"))) {
      return(normalizePath(cand))
    }
  }
  normalizePath(".")
}
REPO   <- find_repo_root()
DATA   <- file.path(REPO, "data", "analysis")
OUTDIR <- file.path(DATA, "rmd")
FIGDIR <- file.path(REPO, "analysis", "figures", "audit")
dir.create(OUTDIR, recursive = TRUE, showWarnings = FALSE)
dir.create(FIGDIR, recursive = TRUE, showWarnings = FALSE)
cat("Figure artifacts → ", FIGDIR, "\n", sep = "")

AUDIT_CSV <- file.path(DATA, "model_features_audit.csv")
stopifnot(file.exists(AUDIT_CSV))
```

```{r helpers}
prepare_audit <- function(dat) {
  dat$question          <- factor(dat$block_id)
  dat$evidence_id       <- factor(dat$evidence_id)
  dat$persona           <- factor(dat$persona)
  dat$agent             <- factor(dat$agent)
  dat$evidence_in_domain <- factor(
    dat$evidence_in_domain,
    levels = c(TRUE, FALSE),
    labels = c("in_domain", "out_of_domain")
  )
  dat$evidence_placement <- factor(dat$evidence_placement)
  dat$topic_drift        <- factor(dat$topic_drift)
  dat$evidence_density   <- factor(dat$evidence_density)
  dat$actual_turns       <- as.numeric(dat$actual_turns)
  dat$retrieved_strict   <- as.integer(dat$retrieved_strict)
  dat
}

tidy_fixef <- function(fit) {
  broom.mixed::tidy(fit, effects = "fixed", conf.int = TRUE)
}

save_csv <- function(x, name) {
  readr::write_csv(as.data.frame(x), file.path(OUTDIR, name))
  invisible(x)
}

anova_type3 <- function(fit, file, caption) {
  tab <- car::Anova(fit, type = 3)
  tab <- tab[rownames(tab) != "(Intercept)", , drop = FALSE]
  df <- as.data.frame(tab)
  df$Term <- rownames(df)
  df <- df[, c("Term", setdiff(names(df), "Term"))]
  save_csv(df, file)
  print(tab)
  knitr::kable(df, digits = 3, row.names = FALSE, caption = caption)
}

save_plot <- function(p, name, width = 8, height = 5, dpi = 300) {
  png_path <- file.path(FIGDIR, paste0(name, ".png"))
  pdf_path <- file.path(FIGDIR, paste0(name, ".pdf"))
  ggplot2::ggsave(png_path, p, width = width, height = height, dpi = dpi, bg = "white")
  ggplot2::ggsave(pdf_path, p, width = width, height = height, bg = "white")
  cat("Wrote ", png_path, "\n", sep = "")
  p
}

audit_contrasts <- list(
  evidence_placement = "contr.sum",
  topic_drift = "contr.sum",
  evidence_density = "contr.sum",
  agent = "contr.sum",
  persona = "contr.sum",
  evidence_in_domain = "contr.sum"
)

fit_audit <- function(dat) {
  glmer(
    retrieved_strict ~ evidence_placement + topic_drift + evidence_density +
      actual_turns + agent + persona + evidence_in_domain +
      (1 | question) + (1 | evidence_id),
    data = dat, family = binomial("logit"),
    contrasts = audit_contrasts, control = glmer_ctrl
  )
}
```

# Data

```{r load-data}
raw <- read_csv(AUDIT_CSV, show_col_types = FALSE)

n_incomplete_rows <- sum(raw$all_required_present %in% FALSE, na.rm = TRUE)
n_incomplete_sess <- raw %>%
  filter(all_required_present %in% FALSE) %>%
  distinct(session_id) %>%
  nrow()
cat(sprintf("Audit rows: %d | unique sessions: %d\n",
            nrow(raw), n_distinct(raw$session_id)))
cat(sprintf("Incomplete (audited missing key information): %d rows / %d sessions\n",
            n_incomplete_rows, n_incomplete_sess))

dat_all <- raw %>%
  filter(!is.na(actual_turns),
         !is.na(evidence_placement),
         !is.na(topic_drift),
         !is.na(evidence_density),
         !is.na(evidence_id),
         !is.na(evidence_in_domain)) %>%
  prepare_audit()

dat_complete <- dat_all %>%
  filter(is.na(all_required_present) | all_required_present)

cat(sprintf("Model sample (all): %d rows\n", nrow(dat_all)))
cat(sprintf("Model sample (complete only): %d rows\n", nrow(dat_complete)))
print(table(dat_all$evidence_in_domain, useNA = "ifany"))
```

```{r describe-incomplete}
raw %>%
  filter(all_required_present %in% FALSE) %>%
  distinct(session_id, evidence_id, persona, evidence_placement,
           topic_drift, evidence_density) %>%
  kable(caption = "Incomplete gold sessions (all_required_present = FALSE).")
```

# Primary audit model (all sessions)

```{r fit-all}
m_all <- fit_audit(dat_all)
print(summary(m_all), correlation = FALSE)
anova_type3(
  m_all,
  "audit_anova_full.csv",
  "Type III Wald tests for the full audit sample (additive model)."
)
```

```{r emmeans-all}
emm_drift <- emmeans(m_all, ~ topic_drift, type = "response")
emm_drift_agent <- emmeans(m_all, ~ topic_drift | agent, type = "response")
emm_persona <- emmeans(m_all, ~ persona, type = "response")
emm_domain <- emmeans(m_all, ~ evidence_in_domain, type = "response")
save_csv(summary(emm_drift), "audit_emmeans_drift.csv")
save_csv(summary(emm_drift_agent), "audit_emmeans_drift_by_agent.csv")
save_csv(summary(emm_persona), "audit_emmeans_persona.csv")
save_csv(summary(emm_domain), "audit_emmeans_in_domain.csv")
print(summary(emm_drift))
print(summary(emm_persona))
print(summary(emm_domain))

fe_all <- tidy_fixef(m_all) %>%
  filter(term == "actual_turns")
kable(fe_all, digits = 3, caption = "Slope of actual_turns (full audit sample).")
```

```{r plot-drift-agent, fig.width=9, fig.height=6, fig.cap="Topic-drift effect on P(strict hit), one panel per agent."}
p_drift <- as.data.frame(summary(emm_drift_agent)) %>%
  mutate(agent = relabel(agent, AGENT_LABELS)) %>%
  ggplot(aes(topic_drift, prob, colour = agent, group = 1)) +
  geom_line(linewidth = .7, alpha = .45) +
  geom_pointrange(aes(ymin = asymp.LCL, ymax = asymp.UCL), linewidth = .55) +
  facet_wrap(~ agent, ncol = 2) +
  scale_colour_discrete(guide = "none") +
  scale_y_continuous(labels = scales::percent_format(accuracy = 1),
                     limits = c(0, 1), expand = expansion(mult = c(0, .05))) +
  labs(x = "Topic drift", y = "P(strict hit)") +
  theme_thesis()
save_plot(p_drift, "fig_audit_drift_by_agent", width = 9, height = 6)
```

```{r plot-persona, fig.cap="Estimated P(strict hit) by eval persona."}
p_persona <- as.data.frame(summary(emm_persona)) %>%
  mutate(persona = relabel(persona, PERSONA_LABELS)) %>%
  ggplot(aes(persona, prob, fill = persona)) +
  geom_col(width = .6) +
  geom_errorbar(aes(ymin = asymp.LCL, ymax = asymp.UCL), width = .15) +
  scale_fill_discrete(guide = "none") +
  scale_y_continuous(labels = scales::percent_format(accuracy = 1),
                     limits = c(0, 1), expand = expansion(mult = c(0, .05))) +
  labs(x = "Persona", y = "P(strict hit)") +
  theme_thesis()
save_plot(p_persona, "fig_audit_persona", width = 6, height = 4)
```

```{r plot-in-domain, fig.cap="Estimated P(strict hit) by whether this gold session is in-domain for the eval persona."}
p_domain <- as.data.frame(summary(emm_domain)) %>%
  mutate(evidence_in_domain = dplyr::recode(
    evidence_in_domain,
    in_domain = "In-domain",
    out_of_domain = "Out-of-domain"
  )) %>%
  ggplot(aes(evidence_in_domain, prob, fill = evidence_in_domain)) +
  geom_col(width = .6) +
  geom_errorbar(aes(ymin = asymp.LCL, ymax = asymp.UCL), width = .15) +
  scale_fill_discrete(guide = "none") +
  scale_y_continuous(labels = scales::percent_format(accuracy = 1),
                     limits = c(0, 1), expand = expansion(mult = c(0, .05))) +
  labs(x = "Gold session vs. eval persona", y = "P(strict hit)") +
  theme_thesis()
save_plot(p_domain, "fig_audit_in_domain", width = 5.5, height = 4)
```

# Sensitivity: drop incomplete sessions

```{r fit-complete}
m_comp <- fit_audit(dat_complete)
anova_type3(
  m_comp,
  "audit_anova_complete.csv",
  "Type III Wald tests after dropping incomplete sessions."
)
```

```{r compare-coefs}
focal_pattern <- paste0(
  "^(evidence_placement|topic_drift|evidence_density|actual_turns)"
)
cmp <- full_join(
  tidy_fixef(m_all) %>%
    filter(grepl(focal_pattern, term)) %>%
    select(term, beta_all = estimate, se_all = std.error),
  tidy_fixef(m_comp) %>%
    filter(grepl(focal_pattern, term)) %>%
    select(term, beta_complete = estimate, se_complete = std.error),
  by = "term"
) %>%
  mutate(delta = beta_complete - beta_all)
save_csv(cmp, "audit_diversity_full_vs_complete.csv")
kable(cmp, digits = 3,
      caption = "Generation-attribute coefficients: all sessions vs. incomplete dropped.")
```

```{r session-info}
sessionInfo()
```
